---
title: "Cities in Higher-HDI Countries Have More Urban Greenspace in South Asia"
author: "Sonika Mahat"
institute: "American University, SIS-750: Data Analysis"
date: "2025-04-09"
output: 
  beamer_presentation:
    theme: "Malmoe"
    colortheme: "beaver"
  fontsize: 12pt
  classoption: "aspectratio=169"
---
```{r setup, include=FALSE}
#Packages
library(knitr)
library(rmdformats)
library(tidyverse)
#Global options
knitr::opts_chunk$set(
  echo = FALSE,
  prompt = FALSE,
  tidy = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center')
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

