---
title: "Cities in Higher-HDI Countries Have More Urban Greenspace in South Asia"
author: "Sonika Mahat"
institute: "American University, SIS-750: Data Analysis"
date: "2025-04-09"
output: 
  beamer_presentation:
    theme: "Berlin"
    colortheme: "whale"
  fontsize: 11pt
  classoption: "aspectratio=169"
---
```{r setup, include=FALSE}
#Packages
library(knitr)
library(rmdformats)
library(tidyverse)
#Global options
knitr::opts_chunk$set(
  echo = FALSE,
  prompt = FALSE,
  tidy = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center')
```
#  Overview: Motivation and Key Findings

##  Motivation

**Why is Greenspace Important in Urban Areas?**

-  Improves air quality, cools cities, and supports biodiversity.  
-  Encourages physical activity, reduces stress, and fosters community.

Urban greenspace in South Asia is under pressure from rapid population growth and limited urban planning, especially in lower-HDI countries.  

-  Note: *HDI = Human Development Index, South Asian countries: India, Nepal, Bangladesh, Pakistan, Sri Lanka, Bhutan, Maldives, Afghanistan.*

##  Research Question and Key Findings

**Research Question:**  

-  How does urban greenspace vary across South Asian countries?
-  Do more developed countries (higher HDI) have greener cities in South Asia?

**Key Findings:**  

-  Cities in higher-HDI countries have significantly more greenspace than those in lower-HDI contexts.

# Dataset: Global Greenspace Indicator

## Global Greenspace Indicator Dataset  

**Data Source:**  

-  Harvard dataverse (2023), NDVI data from 1,000 cities worldwide (2010-2021). *(NDVI = Normalized Difference Vegetation Index)*

**Scope for this Project:**  

-  Cities in South Asia: India, Nepal, Bangladesh, Pakistan, Sri Lanka, Bhutan, Maldives, Afghanistan.

**Key variables used:**  

-  `annual_avg_2021`: NDVI score indicating level of urban greenspace in 2021.
-  `Country`, `City`, `HDI_level`: Compare cities by HDI level.

# Analysis of Dataset: Analysis 1 and 2

## Analysis 1: Greenspace by Country

**Research Question:** How does urban greenspace vary across South Asian countries?

**Method:** Calculated the average NDVI(2021) for cities within each South Asian countries.

## Analysis 1: Greenspace by Country (cont.)

**Visualization:**
```{r analysis1, fig.width=6, fig.height=3.5, out.width="0.8\\linewidth"}
# Load packages
library(tidyverse)

# Read the CSV file
greenspace_data <- read.csv("greenspace_data_share.csv", encoding = "latin1")

# Define South Asian countries
south_asia <- c("Afghanistan", "Bangladesh", "Bhutan", "India", 
                "Maldives", "Nepal", "Pakistan", "Sri Lanka")

# Filter to South Asian countries only
greenspace_sa <- greenspace_data %>%
  filter(Country %in% south_asia)

# Remove rows with missing 2021 NDVI values
greenspace_sa <- greenspace_sa %>%
  filter(!is.na(annual_avg_2021))

# Calculate average NDVI for each country (2021)
country_avg <- greenspace_sa %>%
  group_by(Country) %>%
  summarise(mean_ndvi_2021 = mean(annual_avg_2021, na.rm = TRUE)) %>%
  arrange(desc(mean_ndvi_2021))

# Bar chart using average NDVI in 2021 by country
ggplot(country_avg, aes(x = reorder(Country, -mean_ndvi_2021), 
                        y = mean_ndvi_2021, 
                        fill = mean_ndvi_2021)) + 
  geom_bar(stat = "identity", width = 0.6, show.legend = FALSE) +
  scale_fill_gradient(low = "#c7e9c0", high = "#006d2c") +
  expand_limits(y = 0) +
  labs(
    title = "Average Urban Greenspace by Country (2021)",
    x = "Country",
    y = "Mean NDVI (2021)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5, size = 14),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10))
  )
```
**Takeaway:** Sri Lanka most green, Afghanistan and Maldives least.

## Analysis 2: Greenspace by HDI Level  

**Research Question:** Do cities in more developed South Asian countries with higher HDI have more urban greenspace?

**Method:** Grouped cities by HDI level and compared average NDVI (2021) using a boxplot.

## Analysis 2: Greenspace by HDI Level (cont.)

**Visualization:**
```{r analysis2, fig.width=6, fig.height=3.5, out.width="0.8\\linewidth"}
ggplot(greenspace_sa, aes(x = HDI_level, y = annual_avg_2021, fill = HDI_level)) +
  geom_boxplot(show.legend = FALSE) +
  scale_fill_manual(
    values = c(
      "Low" = "#d9f0d3",      
      "Medium" = "#74c476", 
      "High" = "#006d2c"      
    )
  ) +
  labs(
    title = "Urban Greenspace by HDI Level (2021)",
    x = "HDI Level",
    y = "NDVI Score (2021)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )
```
**Takeaway:** Higher-HDI countries are greener than lower-HDI.

## Analysis 2: Statistical Test- ANOVA (cont.)

**Purpose:** Test whether the differences in urban greenspace across HDI levels are statistically significant.

**Method:** Used one-way ANOVA (Analysis of Variance) to compare NDVI scores (2021) across HDI categories.

## Analysis 2: Statistical Test- ANOVA (cont.)
**Result:**  
```{r ANOVA, results='asis', echo=FALSE}
anova_result <- aov(annual_avg_2021 ~ HDI_level, data = greenspace_sa)
summary(anova_result)
```

**Interpretation:**    
There is a statistically significant difference in greenspace between HDI groups. Cities in higher-HDI countries tend to have significantly greener urban environments.

##  Conclusion  

-  Urban greenspace varies widely across South Asia.  
-  Cities in higher-HDI countries are significantly greener.  
-  ANOVA confirms differences are statistically significant (p < 0.001).
-  Development level plays a key role in urban environmental quality.

#  Thank you